### 2025-05-10 appendices/arima2/appendixArima2.Rnw
Added appendix files, fixed citations, etc.

ggplot(df_combined, aes(x = Re, y = Im, col = Re^2 + Im^2 >= 1, shape = Re^2 + Im^2 >= 1)) +
  geom_point() +
  theme_bw() +
  annotate("path", x = cos(seq(0, 2*pi, length.out = 100)),
           y = sin(seq(0, 2*pi, length.out = 100))) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  coord_fixed() +
  facet_wrap(~alg) +
  scale_color_manual(values = c("black", 'red')) +
  scale_shape_manual(values = c(20, 4)) +
  theme(legend.position = 'none', axis.title = element_blank())
@
\caption{\label{fig:AR2}Inverted roots of 1000 samples of AR(2) and AR(3) coefficients sampled independently from a U$(-1, 1)$ distribution. The red ``x"s are points that lie outside the accepted region, which represents $\Sexpr{round(100*sum((real_parts_AR2^2 + imag_parts_AR2^2 >= 1)) / length(real_parts_AR2), 1)}\%$ of the AR(2) coefficients and $\Sexpr{round(100*sum((real_parts_AR3^2 + imag_parts_AR3^2 >= 1)) / length(real_parts_AR3), 1)}\%$ of the AR(3) coefficients. Increasing the width of the uniform sampling distribution results in a larger fraction outside the accepted region, and decreasing the width results in worse coverage of the range of acceptable parameter values.}
\end{figure}

>>>>

ggplot(df_combined, aes(x = Re, y = Im, col = Re^2 + Im^2 >= 1, shape = Re^2 + Im^2 >= 1)) +
  geom_point() +
  theme_bw() +
  annotate("path", x = cos(seq(0, 2*pi, length.out = 100)),
           y = sin(seq(0, 2*pi, length.out = 100))) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  coord_fixed() +
  facet_wrap(~alg) +
  scale_color_manual(values = c("black", 'red')) +
  scale_shape_manual(values = c(20, 4)) +
  theme(legend.position = 'none', axis.title = element_blank())
@
\caption[Uniform resampling of model parameters.]{\label{fig:AR2}Inverted roots of 1000 samples of AR(2) and AR(3) coefficients sampled independently from a U$(-1, 1)$ distribution. The red ``x"s are points that lie outside the accepted region, which represents $\Sexpr{round(100*sum((real_parts_AR2^2 + imag_parts_AR2^2 >= 1)) / length(real_parts_AR2), 1)}\%$ of the AR(2) coefficients and $\Sexpr{round(100*sum((real_parts_AR3^2 + imag_parts_AR3^2 >= 1)) / length(real_parts_AR3), 1)}\%$ of the AR(3) coefficients. Increasing the width of the uniform sampling distribution results in a larger fraction outside the accepted region, and decreasing the width results in worse coverage of the range of acceptable parameter values.}
\end{figure}

### 2025-05-10 chapters/arima2/arima2.Rnw
Added appendix files, fixed citations, etc.

\begin{figure}[ht]
<<mutliMode-fig, fig.height=3.5, fig.width=7>>=
ggpubr::ggarrange(
  plotlist = ggs,
  legend = 'right',
  common.legend = TRUE
)
@
\caption{\label{fig:multiMode}The profile log-likelihood of data simulated from four distinct MA(1) models, demonstrating a few examples of multimodal likelihood surfaces. The solid, black line indicates the true value of $\theta_1$; the dotted line is the CSS-initialization. The dashed lines correspond to the estimate $\hat{\theta}_1$ using \code{stats:arima} (red) and our proposed algorithm (implemented in \code{arima2::arima}, blue).}
\end{figure}

@@

\begin{figure}[!ht]
<<simProps, fig.height=3, fig.width=4.5>>=
wide_res %>%
  filter(P != 0 & Q !=0) %>%
  group_by(P, Q, n) %>%
  summarize(pct_improve = sum(ll_arima2 > ll_arima, na.rm = TRUE) / n()) %>%
  ggplot(aes(x = n, y = pct_improve)) +
  facet_grid(Q~P, labeller = labeller(P = AR_labs, Q = MA_labs)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(limits = c(0, 0.6)) +
  ylab("Proportion of Improved Fits") +
  xlab("Number of observations (n)")
@
\caption{\label{fig:simProps}Proportion of simulated data with improved likelihood from using multiple restarts (Algorithm~\ref{alg:mle}).}
\end{figure}

@@

ggpubr::ggarrange(
  gg_fisher, gg_profile, align = 'h', nrow = 1,
  labels = 'AUTO',
  label.x = c(0.16, 0)
)
@
\caption{\label{fig:pctCovered}Proportion of models that achieved nominal coverage of Bonferroni adjusted $95\%$ confidence intervals. The dashed line denotes the target coverage level. (A) Confidence intervals created using Fisher's information matrix. (B) Confidence intervals created using profile likelihoods.}
\end{figure}

@@

simTables %>%
  group_by(n, P, Q, max_repeat) %>%
  summarize(
    pct_consistent = sum(consistent, na.rm = TRUE) / sum(!is.na(consistent))
  ) %>%
  ggplot(
    aes(
      x = n, y = pct_consistent,
      group = as.factor(max_repeat),
      color = as.factor(max_repeat),
      linetype = as.factor(max_repeat)
    )
  ) +
  geom_line() +
  geom_point() +
  facet_grid(P~Q, labeller = labeller(P = AR_labs, Q = MA_labs)) +
  ylab("Proportion of Tables\n that are Consistent") +
  xlab("Number of observations") +
  theme(legend.title = element_blank(), legend.position = 'bottom') +
  scale_color_discrete(labels = restart_labels) +
  scale_linetype_discrete(labels = restart_labels) +
  scale_y_continuous(limits = c(0.35, 1))
@
\caption{\label{fig:AICtabRes}Data is generated from ARMA$(p, q)$ models with $(p, q) \in \{1, 2, 3\}^2$, and the corresponding AIC table is created. The Y-axis shows the percentage of tables that were consistent. M is the number of times a maxima is observed before the algorithm terminates, so $M=1$ corresponds to the standard maximization procedure.}
\end{figure}

@@

    \caption{\label{tab:huronTab}AIC values for an ARMA(p, q) model fit to Lake Michigan-Huron depths. Table~\ref{tab:AIC1} was computing using only a single parameter initialization. Table~\ref{tab:AIC2} was computed using Algorithm~\ref{alg:mle}. Highlighted cells show where the likelihood was improved (AIC reduced) using our algorithm.}

@@

\begin{table}
\centering
\caption{\label{tab:huronCoefsTab}Parameter values of $\text{ARMA}(p, q)$ model fit to Lake Michigan-Huron depth data.}
\begin{tabular}{c|c|c|c|c}\hline
 & $\phi_1$ & $\phi_2$ & $\theta_1$ & Intercept \\\hline
 Estimate & \Sexpr{format(round(huron_21$coef['ar1'], 3), nsmall = 3)} & \Sexpr{format(round(huron_21$coef['ar2'], 3), nsmall = 3)} & \Sexpr{format(round(huron_21$coef['ma1'], 3), nsmall = 3)}  & \Sexpr{format(round(huron_21$coef['intercept'], 3), nsmall = 3)} \\\hline
 s.e. & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['ar1'], 3), nsmall = 3)} & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['ar2'], 3), nsmall = 3)} & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['ma1'], 3), nsmall = 3)}  & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['intercept'], 3), nsmall = 3)} \\\hline
\end{tabular}
\end{table}

@@

cowplot::plot_grid(gg_prof, gg_ma21, gg_ma1, nrow = 1, labels = 'AUTO')
@
\caption{\label{fig:huronMAevid}Evidence for an AR(1) model for the Lake Michigan-Huron data. (A) Profile likelihood confidence interval (PLCI) for $\theta_1$ which includes the value $\theta_1 = 0$. The vertical dotted line represents the lower end of the approximate confidence interval; all points on the solid black line lie within the confidence interval, and points on the dashed red line are outside the interval. (B) Histogram of re-estimated $\theta_1$ values using simulated data simulated from the $\text{ARMA}(2, 1)$ model that was calibrated to the Lake Michigan-Huron data. (C) Histogram of re-estimated $\theta_1$ values using data simulated from the AR(1) model that was calibrated to the Lake Michigan-Huron data.}
\end{figure}

@@

ggplot(all_roots, aes(x = real, y = im, col = mod, pch = mod)) +
  annotate("path", x = cos(seq(0, 2*pi, length.out = 100)),
           y = sin(seq(0, 2*pi, length.out = 100))) +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  geom_vline(xintercept = 0, linewidth = 0.3) +
  geom_point(aes(size = mod), stroke = 1.2) +
  facet_wrap(
    ~type,
    nrow = 1,
    labeller = as_labeller(
      c("AR" = "Inverse AR roots", "MA" = "Inverse MA roots"))
  ) +
  coord_fixed() +
  labs(
    color = NULL, pch = NULL, stroke = NULL, size = NULL, y = 'Imaginary', x = 'Real'
    ) +
  theme_bw() +
  theme(axis.title = ggplot2::element_text(face="bold"),
        legend.position = 'bottom') +
  scale_shape_manual(values = c(1, 4, 3)) +
  scale_size_manual(values = c(3.7, 2.7, 2.7)) +
  scale_color_brewer(palette = 'Dark2')
@
\caption{\label{fig:huronRoots}Inverted AR and MA polynomial roots to the fitted $\text{AR}(1)$ and $\text{ARMA}(2, 1)$ models to the Lake Michigan-Huron data using a single parameter initialization.}
\end{figure}

>>>>

\begin{figure}[ht]
<<mutliMode-fig, fig.height=3.5, fig.width=7>>=
ggpubr::ggarrange(
  plotlist = ggs,
  legend = 'right',
  common.legend = TRUE
)
@
\caption[Multi-node MA(1) models]{\label{fig:multiMode}The profile log-likelihood of data simulated from four distinct MA(1) models, demonstrating a few examples of multimodal likelihood surfaces. The solid, black line indicates the true value of $\theta_1$; the dotted line is the CSS-initialization. The dashed lines correspond to the estimate $\hat{\theta}_1$ using \code{stats:arima} (red) and our proposed algorithm (implemented in \code{arima2::arima}, blue).}
\end{figure}

@@

\begin{figure}[!ht]
<<simProps, fig.height=3, fig.width=4.5>>=
wide_res %>%
  filter(P != 0 & Q !=0) %>%
  group_by(P, Q, n) %>%
  summarize(pct_improve = sum(ll_arima2 > ll_arima, na.rm = TRUE) / n()) %>%
  ggplot(aes(x = n, y = pct_improve)) +
  facet_grid(Q~P, labeller = labeller(P = AR_labs, Q = MA_labs)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(limits = c(0, 0.6)) +
  ylab("Proportion of Improved Fits") +
  xlab("Number of observations (n)")
@
\caption[Proportion of improved models]{\label{fig:simProps}Proportion of simulated data with improved likelihood from using multiple restarts (Algorithm~\ref{alg:mle}).}
\end{figure}

@@

ggpubr::ggarrange(
  gg_fisher, gg_profile, align = 'h', nrow = 1,
  labels = 'AUTO',
  label.x = c(0.16, 0)
)
@
\caption[Confidence interval coverage]{\label{fig:pctCovered}Proportion of models that achieved nominal coverage of Bonferroni adjusted $95\%$ confidence intervals. The dashed line denotes the target coverage level. (A) Confidence intervals created using Fisher's information matrix. (B) Confidence intervals created using profile likelihoods.}
\end{figure}

@@

simTables %>%
  group_by(n, P, Q, max_repeat) %>%
  summarize(
    pct_consistent = sum(consistent, na.rm = TRUE) / sum(!is.na(consistent))
  ) %>%
  ggplot(
    aes(
      x = n, y = pct_consistent,
      group = as.factor(max_repeat),
      color = as.factor(max_repeat),
      linetype = as.factor(max_repeat)
    )
  ) +
  geom_line() +
  geom_point() +
  facet_grid(P~Q, labeller = labeller(P = AR_labs, Q = MA_labs)) +
  ylab("Proportion of Tables\n that are Consistent") +
  xlab("Number of observations") +
  theme(legend.title = element_blank(), legend.position = 'bottom') +
  scale_color_discrete(labels = restart_labels) +
  scale_linetype_discrete(labels = restart_labels) +
  scale_y_continuous(limits = c(0.35, 1))
@
\caption[AIC table improvements]{\label{fig:AICtabRes}Data is generated from ARMA$(p, q)$ models with $(p, q) \in \{1, 2, 3\}^2$, and the corresponding AIC table is created. The Y-axis shows the percentage of tables that were consistent. M is the number of times a maxima is observed before the algorithm terminates, so $M=1$ corresponds to the standard maximization procedure.}
\end{figure}

@@

    \caption[AIC values of ARMA model for Lake Michigan-Huron]{\label{tab:huronTab}AIC values for an ARMA(p, q) model fit to Lake Michigan-Huron depths. Table~\ref{tab:AIC1} was computing using only a single parameter initialization. Table~\ref{tab:AIC2} was computed using Algorithm~\ref{alg:mle}. Highlighted cells show where the likelihood was improved (AIC reduced) using our algorithm.}

@@

\begin{table}
\centering
\caption{\label{tab:huronCoefsTab}Parameter values of an ARMA model fit to Lake Michigan-Huron data.}
\begin{tabular}{c|c|c|c|c}\hline
 & $\phi_1$ & $\phi_2$ & $\theta_1$ & Intercept \\\hline
 Estimate & \Sexpr{format(round(huron_21$coef['ar1'], 3), nsmall = 3)} & \Sexpr{format(round(huron_21$coef['ar2'], 3), nsmall = 3)} & \Sexpr{format(round(huron_21$coef['ma1'], 3), nsmall = 3)}  & \Sexpr{format(round(huron_21$coef['intercept'], 3), nsmall = 3)} \\\hline
 s.e. & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['ar1'], 3), nsmall = 3)} & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['ar2'], 3), nsmall = 3)} & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['ma1'], 3), nsmall = 3)}  & \Sexpr{format(round(sqrt(diag(huron_21$var.coef))['intercept'], 3), nsmall = 3)} \\\hline
\end{tabular}
\end{table}

@@

cowplot::plot_grid(gg_prof, gg_ma21, gg_ma1, nrow = 1, labels = 'AUTO')
@
\caption[Evidence for an AR(1) model for the Lake Michigan-Huron data.]{\label{fig:huronMAevid}Evidence for an AR(1) model for the Lake Michigan-Huron data. (A) Profile likelihood confidence interval (PLCI) for $\theta_1$ which includes the value $\theta_1 = 0$. The vertical dotted line represents the lower end of the approximate confidence interval; all points on the solid black line lie within the confidence interval, and points on the dashed red line are outside the interval. (B) Histogram of re-estimated $\theta_1$ values using simulated data simulated from the $\text{ARMA}(2, 1)$ model that was calibrated to the Lake Michigan-Huron data. (C) Histogram of re-estimated $\theta_1$ values using data simulated from the AR(1) model that was calibrated to the Lake Michigan-Huron data.}
\end{figure}

@@

ggplot(all_roots, aes(x = real, y = im, col = mod, pch = mod)) +
  annotate("path", x = cos(seq(0, 2*pi, length.out = 100)),
           y = sin(seq(0, 2*pi, length.out = 100))) +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  geom_vline(xintercept = 0, linewidth = 0.3) +
  geom_point(aes(size = mod), stroke = 1.2) +
  facet_wrap(
    ~type,
    nrow = 1,
    labeller = as_labeller(
      c("AR" = "Inverse AR roots", "MA" = "Inverse MA roots"))
  ) +
  coord_fixed() +
  labs(
    color = NULL, pch = NULL, stroke = NULL, size = NULL, y = 'Imaginary', x = 'Real'
    ) +
  theme_bw() +
  theme(axis.title = ggplot2::element_text(face="bold"),
        legend.position = 'bottom') +
  scale_shape_manual(values = c(1, 4, 3)) +
  scale_size_manual(values = c(3.7, 2.7, 2.7)) +
  scale_color_brewer(palette = 'Dark2')
@
\caption[ARMA polynomial roots]{\label{fig:huronRoots}Inverted AR and MA polynomial roots to the fitted $\text{AR}(1)$ and $\text{ARMA}(2, 1)$ models to the Lake Michigan-Huron data using a single parameter initialization.}
\end{figure}
